Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-04-19 SIR_manuscript.tex
typo fixed in a_3

Let define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, thus
\begin{equation}
\label{eq:phi}
\rho=\frac{\omega \phi}{\omega+\phi}.
\end{equation}
Substituding in $\rho$ in the expression of $\Rnum$, in Eq.\,\eqref{R0}, and applying the derivative with respect to $\phi$ results in the following simplified expression for $\partial\Delta/\partial\phi$.

with an equivalent but simplified $\Delta$ in Eq.\,\eqref{eq:abc} as
\begin{equation}
\label{eq:del3}
\Delta= \frac{K_1\,\phi^2+K_2\,\phi}{K_3},
\end{equation}
where
\begin{align}
\label{eq:K}
K_1=& \frac{w_I}{w_S}\Big(((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big), \\
K_2=& (\omega+\gamma)\Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
K_3=& (\omega+\gamma) \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)(\omega+\phi).
\end{align}
The derivative is
\begin{align}
\label{eq:dd3dr}
\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3\,\phi^2+b_3\,\phi+c_3),
\end{align}
where
\begin{align}
\label{eq:abcd2}
a_3=& \frac{w_I}{w_S} \Big( (1+\gamma)(\theta\_w-\theta\_c)\frac{w_I}{w_S}\omega\,p_I+ (\gamma\theta\_w+\omega\,p_I\theta\_c)(\omega+\gamma) \Big),\\
b_3=& 2\frac{w_I}{w_S}(\omega+\gamma)\Big( ((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big) ,\\
c_3=& (\omega+\gamma)^2 \Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
d_3=& \frac{(\omega+\gamma)}{\omega\gamma} \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2. 
\end{align}

Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$. 
If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$. An example of this scenario occures when $\theta\_w=0$ and $\theta\_c=1$, thus $\theta\_w-\theta\_c=-1$, this is when the strength of isolation for awaiting people is the least, but the most for the confirmed cases. In this case 
\begin{equation}
\label{ineq:a3}
a_3 <0 \iff \omega+\gamma<(1+\gamma)\frac{w_I}{w_S}.
\end{equation}
Given that $\frac{w_I}{w_S}\geq 1$, the sufficient condition for this inequality to be held is 
$\omega < \frac{w_I}{w_S}$.
This can be visualized in the top-right panel of the Fig.\,\ref{pan2} panel (b).

>>>>

Let define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, thus
\begin{equation}
\label{eq:phi}
\rho=\frac{\omega \phi}{\omega+\phi}.
\end{equation}
Reparametrizing testing rate $\rho$ by $\phi$ in the expression of $\Rnum$, in Eq.\,\eqref{R0}, and applying the derivative with respect to $\phi$ results in the following simplified expression for $\partial\Delta/\partial\phi$.

with an equivalent but simplified $\Delta$ in Eq.\,\eqref{eq:abc} as
\begin{equation}
\label{eq:del3}
\Delta= \frac{K_1\,\phi^2+K_2\,\phi}{K_3},
\end{equation}
where
\begin{align}
\label{eq:K}
K_1=& \frac{w_I}{w_S}\Big(((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big), \\
K_2=& (\omega+\gamma)\Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
K_3=& (\omega+\gamma) \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)(\omega+\phi).
\end{align}
The derivative is
\begin{align}
\label{eq:dd3dr}
\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3\,\phi^2+b_3\,\phi+c_3),
\end{align}
where
\begin{align}
\label{eq:abcd2}
a_3=& \frac{w_I}{w_S} \Big((\omega\,p_I+\gamma)(\theta\_w-\theta\_c)\frac{w_I}{w_S}\omega\,p_I+(\gamma\theta\_w+\omega\,p_I\theta\_c)(\omega+\gamma) \Big),\\
b_3=& 2\frac{w_I}{w_S}(\omega+\gamma)\Big( ((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big) ,\\
c_3=& (\omega+\gamma)^2 \Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
d_3=& \frac{(\omega+\gamma)}{\omega\gamma} \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2. 
\end{align}

Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Delta \leq 0$. 
If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$. An example of this scenario occures when $\theta\_w=0$ and $\theta\_c=1$, which is illustraited in the top-right panel of the Fig.\,\ref{pan2} panel (b). This is when the strength of isolation for awaiting people is the least, but the most for the confirmed cases. In this case 
\begin{equation}
\label{ineq:a3}
a_3 <0\iff\omega+\gamma<(\omega\,p_I+\gamma)\frac{w_I}{w_S}.
\end{equation}
Given that $\frac{w_I}{w_S}\geq 1$, and $0\leq\,p_I\leq\,1$, this inequality is hold when a perfect sensitive test is used, i.e., $p_I=1$, and a targeted tesing strategy is followed, i.e., $\frac{w_I}{w_S}>1$. \ali{i)what is the interpretation of this? 2) (Ben:) calculate the corresponding eigenvector to $\Rnum$, scale it to 1, and plot the population of each infected compartment, what does this tell us? }

Note that $\rho$ and $\omega$ have similar mechanism in delaying people to get into $I\_c$. The non monotonic countereffect of testing rate and test return rate are not obvious. 
that people in the DFE waiting for negative test, $S_n^*$.
More testing may have a counter effect on $\Rnum$, as if we have more testing we have more people in $S_n^*$ which will take them longer to get to $I\_c$.

more $\rho$ then
